\begin{abstract}
Retrieval-augmented generation (RAG) and graph-based retrieval (GraphRAG)
are the dominant paradigms for grounding LLM responses in structured
knowledge. Both optimize for recall while treating token cost as a
secondary concern. We introduce Compact Knowledge Graphs (CKG)---pre-structured
DAG representations with explicit concept taxonomy and pipe-delimited
dependency encoding---and present a two-track benchmark comparing CKG,
RAG, and GraphRAG across educational and commercial domains.

Track 1 evaluates the McCreary Intelligent Textbook Corpus: 44
hand-curated educational domains spanning STEM, professional, and
foundational subjects (7,758 queries). CKG achieves macro-average
F1 of \textbf{0.4709} versus \textbf{0.1231} for RAG and \textbf{0.1200}
for GraphRAG, at \textbf{11$\times$ fewer tokens per query} (269 vs.\
2,982 vs.\ 3,450). CKG F1 increases with hop depth (peaking at hop=3:
0.64) while RAG degrades. The compound Reasoning Density Score (RDS)
advantage is \textbf{42$\times$}. T1 entity lookup (CKG 0.207, RAG 0.094)
serves as a designed negative control.

Track 2 tests whether the architecture generalizes beyond hand-curated
educational data. We build a complete GLP-1/Obesity pharmacology CKG
programmatically from the ClinicalTrials.gov API with no expert
annotation (170 queries). The pipeline-generated CKG reaches macro F1
\textbf{0.5298}---exceeding the Track 1 hand-curated average by 12.5\%
and preserving a \textbf{28$\times$} RDS advantage over RAG. Automated
construction does not degrade retrieval quality; it matches or improves
it. Together the two tracks establish CKG as a distinct architecture
whose structural advantage is domain-agnostic and does not depend on
manual curation.

Benchmark, dataset, and evaluation harness are released at
\url{https://huggingface.co/datasets/graphify-md/ckg-benchmark}
under CC BY 4.0\,/\,MIT.
\end{abstract}
