\section{Metrics}\label{sec:metrics}

We evaluate 16 metrics organized into six categories. Novel metrics introduced
in this paper are marked with~$\star$.

\subsection{Standard IR Metrics}\label{sec:metrics-standard}

\paragraph{Token-Level F1 (SQuAD-style).}
Used for T1, T2, and T4 queries. Precision, recall, and F1 are computed over
token sets:
\begin{equation}
  \text{F1} = \frac{2 \cdot P \cdot R}{P + R}, \quad
  P = \frac{|pred \cap truth|}{|pred|}, \quad
  R = \frac{|pred \cap truth|}{|truth|}
\end{equation}

\paragraph{Edge-Overlap F1.}
Used for T3 (path) and T5 (cross-concept) queries:
\begin{equation}
  \text{Edge\_F1} = \frac{2 \cdot |E_{pred} \cap E_{truth}|}{|E_{pred}| + |E_{truth}|}
\end{equation}

\paragraph{Exact Match (EM).}
Binary---the full answer must match ground truth exactly. Reported alongside F1
as a secondary metric.

\subsection{Reasoning Density Score$^\star$ (RDS)}\label{sec:metrics-rds}

The core compound metric introduced in this paper:
\begin{equation}\label{eq:rds}
  \text{RDS}(s, q) = \frac{\text{F1}(s, q)}{\text{tokens\_consumed}(s, q)}
\end{equation}

Macro-averaged across all queries:
$\text{RDS}_{\text{macro}}(s) = \text{mean}(\text{RDS}$ over all queries$)$.
The \textbf{RDS ratio} compares two systems:
$\text{RDS}_{\text{ratio}}(A, B) = \text{RDS}_{\text{macro}}(A) / \text{RDS}_{\text{macro}}(B)$.
Higher values indicate more reasoning quality per token spent.

\subsection{Hop-Depth F1 Degradation$^\star$}\label{sec:metrics-hop}

Measures how F1 degrades as reasoning chain length increases:
\begin{equation}
  \text{F1@hop}(s, k) = \text{mean}(\text{F1} \mid \text{hop\_depth} = k)
\end{equation}

Reported for $k = 1, 2, 3, 4, 5{+}$. Expected finding: RAG degrades steeply at
$k \geq 2$; CKG remains flat due to explicit edge traversal.

\subsection{Tokenomics Metrics}\label{sec:metrics-token}

\paragraph{Context Utilization Rate$^\star$ (CUR).}
Fraction of retrieved tokens relevant to the answer:
$\text{CUR} = \text{relevant\_tokens} / \text{total\_retrieved\_tokens}$.

\paragraph{Cost Per Correct Answer$^\star$ (CPCA).}
Real-world cost using Claude Sonnet 4.6 pricing (\$3/M input, \$15/M output):
$\text{CPCA} = \text{cost\_per\_query} / \text{F1}$.

\paragraph{Precision at Token Budget (P@T).}
Mean F1 over queries where tokens\_consumed $\leq$ budget $T$.
Reported for $T = 500, 1000, 2000, 5000, 10000$.

\paragraph{Token Budget Breakeven.}
Minimum budget where RAG/GraphRAG F1 $\geq$ CKG F1:
$\text{breakeven} = \min T$ such that $\text{F1}_{\text{RAG}}(T) \geq \text{F1}_{\text{CKG}}(500)$.

\paragraph{Index Build Cost.}
One-time cost: tokens consumed during indexing + wall-clock time + storage.
CKG: zero (CSV already exists).

\paragraph{Update Cost$^\star$.}
Cost to incorporate one new concept: CKG edits one CSV row with zero
re-indexing; RAG re-embeds affected chunks; GraphRAG requires full
re-extraction.

\subsection{Structural Fidelity Metrics}\label{sec:metrics-structural}

\paragraph{Relationship Precision$^\star$ (RP).}
Of edges returned, what fraction are real DAG edges:
$\text{RP} = |E_{pred} \cap E_{truth}| / |E_{pred}|$.

\paragraph{Hub Node Recall$^\star$ (HNR).}
Recall on high-indegree concepts (top 20\% by indegree).

\paragraph{Boundary Completeness$^\star$ (BC).}
For T4 queries, fraction of the taxonomy category returned:
$\text{BC} = |retrieved \cap members| / |members|$.
CKG achieves BC $\approx$ 1.0 by construction.

\subsection{Robustness Metrics}\label{sec:metrics-robustness}

\paragraph{Paraphrase Stability (PS).}
F1 variance across 5 paraphrased versions of each query.
CKG should be stable (exact concept match); RAG is embedding-sensitive.

\paragraph{Hallucination Rate (HR).}
Fraction of queries returning $\geq$1 concept not in the corpus.
CKG: HR $=$ 0 by construction. GraphRAG's dynamic extraction can hallucinate.
