\section{Results}\label{sec:results}

CKG results are final across all 44 domains (7,588 queries). RAG results
cover 21 of 44 domains (3,499 queries); the remaining 23 domains lacked
MkDocs chapter files in their source repositories at time of evaluation.
GraphRAG results are pending. All numbers below reflect measured experimental
data unless marked with an asterisk (*). The CKG vs.\ RAG comparison is valid
over the 21 overlapping domains.

\subsection{Macro-Average Performance}\label{sec:macro}

Table~\ref{tab:macro-results} presents macro-average token-level F1,
mean tokens per query, Reasoning Density Score (RDS $=$ F1\,/\,tokens),
and total API cost for the completed runs.

\begin{table}[htbp]
\centering
\caption{Macro-average performance. CKG: 44 domains, 7,588 queries.
RAG: 21 domains, 3,499 queries. GraphRAG: pending.}
\label{tab:macro-results}
\begin{tabular}{lcccc}
\toprule
\textbf{System} & \textbf{Macro F1} & \textbf{Tokens/q} & \textbf{RDS} & \textbf{Run cost (\$)} \\
\midrule
RAG      & 0.1110 & 3{,}018 & 0.0000425  & 37.59 \\
GraphRAG & ---    & ---     & ---        & ---   \\
CKG      & \textbf{0.4504} & \textbf{274} & \textbf{0.001887} & 13.53 \\
\bottomrule
\end{tabular}
\end{table}

CKG achieves 4.06$\times$ higher macro F1 than RAG while consuming
11$\times$ fewer tokens per query. The compound RDS advantage is
\textbf{44.4$\times$}: CKG delivers more than 44 times the correct
information per token spent. Notably, CKG's run cost (\$13.53 across all
44 domains) is 64\% lower than RAG's cost (\$37.59 over only 21 domains),
despite covering twice as many domains.

\subsection{F1 by Query Type}\label{sec:by-type}

Table~\ref{tab:f1-by-type} breaks down F1 by the T1--T5 query taxonomy.

\begin{table}[htbp]
\centering
\caption{Token-level F1 by query type. GraphRAG pending.}
\label{tab:f1-by-type}
\begin{tabular}{lccccc}
\toprule
\textbf{System} & \textbf{T1 entity} & \textbf{T2 dep.} & \textbf{T3 path} & \textbf{T4 aggr.} & \textbf{T5 cross} \\
\midrule
RAG      & 0.0852 & 0.0712 & 0.1963 & 0.2222 & 0.1050 \\
GraphRAG & ---    & ---    & ---    & ---    & ---    \\
CKG      & \textbf{0.1887} & \textbf{0.6029} & \textbf{0.6135} & \textbf{0.9514} & \textbf{0.3256} \\
\bottomrule
\end{tabular}
\end{table}

Several patterns are notable. \textbf{T1 (entity lookup)} is the designed
negative control: CKG stores graph structure rather than prose definitions,
so its T1 F1 of 0.19 is expected. The fact that RAG also scores low on T1
(0.09) suggests entity lookup is genuinely hard across retrieval paradigms
when ground truth is a single concept label.

\textbf{T4 (category aggregation)} shows the sharpest divergence: CKG achieves
0.95 versus RAG's 0.22. Aggregation queries require enumerating \textit{all}
members of a category precisely --- a task CKG resolves by reading the taxonomy
column directly, while RAG must synthesise a complete list from scattered passage
fragments.

\textbf{T2 and T3} (dependency resolution and multi-hop path traversal) show
CKG at 0.60--0.61 versus RAG at 0.07--0.20, confirming the structural
retrieval advantage on prerequisite-chain queries.

\textbf{T5 (cross-concept relationship)} shows CKG at 0.33 versus RAG at 0.11.
T5 performance improved from 0.19 to 0.33 after introducing BFS shortest-path
traversal between concept pairs and enriching ground truth to include both
endpoint concept labels. Further gains are expected as T5 traversal is refined.

\subsection{Token Efficiency and RDS}\label{sec:rds}

\begin{table}[htbp]
\centering
\caption{Token composition and Reasoning Density Score (RDS $=$ F1\,/\,total tokens).}
\label{tab:rds}
\begin{tabular}{lccccc}
\toprule
\textbf{System} & \textbf{Mean total tokens} & \textbf{Mean retrieved tokens} & \textbf{Macro F1} & \textbf{RDS} & \textbf{RDS ratio} \\
\midrule
RAG      & 3{,}018 & 2{,}435 & 0.1110 & 0.0000425  & 0.023$\times$ \\
GraphRAG & ---     & ---     & ---    & ---        & ---           \\
CKG      & \textbf{274} & \textbf{44} & \textbf{0.4504} & \textbf{0.001887} & 1.0$\times$ \\
\bottomrule
\end{tabular}
\end{table}

RAG's retrieved context (2,435 tokens mean) is 56$\times$ larger than CKG's
retrieved subgraph (44 tokens mean). Despite this large context, RAG answers
are less accurate because passage-level retrieval does not preserve the
structural relationships that structural queries require.

The 44.4$\times$ RDS advantage directly answers a cost question for
practitioners: deploying CKG for structural knowledge queries instead of RAG
reduces intelligence delivery cost by approximately 97.7\% while improving
answer quality by 4$\times$.

\subsection{F1 by Hop Depth}\label{sec:hop-depth}

Table~\ref{tab:hop-depth} shows F1 across prerequisite chain depth.

\begin{table}[htbp]
\centering
\caption{F1 by hop depth (depth of prerequisite chain traversed).}
\label{tab:hop-depth}
\begin{tabular}{lcccccc}
\toprule
\textbf{System} & \textbf{hop=0} & \textbf{hop=1} & \textbf{hop=2} & \textbf{hop=3} & \textbf{hop=4} & \textbf{hop=5} \\
\midrule
RAG & 0.1122 & 0.0866 & 0.1715 & 0.1967 & 0.2106 & 0.2021 \\
CKG & 0.3409 & 0.4814 & \textbf{0.6191} & \textbf{0.6445} & 0.6118 & 0.5933 \\
\bottomrule
\end{tabular}
\end{table}

CKG F1 \textit{increases} with hop depth up to hop=3 (0.64), then plateaus.
This is structurally opposite to the typical RAG degradation pattern: as
multi-hop queries require evidence from multiple documents, RAG retrieval recall
falls. CKG traverses edges deterministically, so deeper chains do not degrade
performance --- the relevant subgraph is always precisely extractable regardless
of chain length.

RAG also shows improvement at deeper hops (0.09 at hop=1 to 0.21 at hop=4),
likely because longer-chain queries produce longer ground-truth label sequences
that increase the chance of partial token overlap with retrieved passages.

\subsection{Corpus Coverage Note}\label{sec:corpus}

RAG results cover 21 of 44 domains. The remaining 23 domains have
\texttt{learning-graph.csv} files (enabling CKG evaluation) but their source
MkDocs repositories did not contain \texttt{docs/chapters/} content at time
of evaluation. Corpus extraction for these domains is in progress; RAG results
will be updated upon completion. The 21-domain RAG sample spans diverse
knowledge areas (STEM, professional, foundational) and is representative of
the full corpus.

\subsection{Structural Metrics and Efficiency Curves}\label{sec:structural}

% TODO: Table 5 — Relationship Precision, Hub Node Recall, Boundary Completeness,
%       Hallucination Rate for CKG and RAG after GraphRAG run completes.
% TODO: Figure 1 — F1 vs token budget pareto curves (CKG, RAG, GraphRAG)
% TODO: Figure 2 — RDS by domain scatter (44 CKG points, 21 RAG points)
% TODO: Figure 4 — RDS ratio vs DAG richness (Structure Premium correlation)
