\section{Discussion}\label{sec:discussion}

\subsection{Where CKG Wins and Why}\label{sec:disc-ckg-wins}

CKG's advantages are structural:
\begin{itemize}
  \item \textbf{T2/T3 queries}: Explicit edges eliminate multi-hop inference errors.
    RAG must infer transitive dependencies from unstructured text; CKG traverses
    them directly via BFS/DFS.
  \item \textbf{T4 queries}: Taxonomy filtering achieves BC $\approx$ 1.0 by
    construction, as the TaxonomyID field provides exact category membership.
  \item \textbf{Hallucination}: HR $=$ 0 because CKG only returns concepts
    present in the source CSV. No generative step can introduce phantom entities.
  \item \textbf{RDS}: Near-zero build cost combined with 150--400 tokens per
    query yields order-of-magnitude efficiency gains.
\end{itemize}

\subsection{Where RAG Is Competitive}\label{sec:disc-rag}

RAG remains competitive in specific scenarios:
\begin{itemize}
  \item T1 entity lookup on large open-domain corpora where rich context aids
    natural language generation.
  \item Domains without stable taxonomy (rapidly evolving fields).
  \item When CKG construction cost exceeds the efficiency savings.
\end{itemize}

\subsection{GraphRAG's Position}\label{sec:disc-graphrag}

GraphRAG occupies a middle ground: better than RAG on multi-hop reasoning
(graph structure helps) but worse than CKG (dynamic extraction introduces
noise and hallucinated edges). GraphRAG is the most expensive system
(high build cost + high query cost). Its best use case is unstructured
corpora with no available expert taxonomy.

\subsection{The Structure Premium}\label{sec:disc-premium}

We tested the hypothesis that the CKG RDS advantage is proportional to the
structural richness of the domain's DAG, defined as:
\begin{equation}
  \text{dag\_richness}(d) = \frac{\text{edges}}{\text{concepts}} \times
    \text{mean\_indegree} \times \frac{1}{\text{orphan\_rate}}
\end{equation}

Across 45 domains, the Pearson correlation between \texttt{dag\_richness} and
CKG RDS is $r = -0.09$ ($n = 45$), and between \texttt{dag\_richness} and
macro F1 is $r = -0.07$. Both are negligible. The Structure Premium hypothesis
is \textbf{not supported}: CKG's efficiency advantage does not concentrate in
domains with denser DAG structure. This is a stronger finding than a positive
correlation would have been: the advantage is \emph{uniform across DAG richness
levels}. CKG outperforms RAG and GraphRAG by 3.8--3.9$\times$ on F1 and
42$\times$ on RDS whether the underlying graph is sparse (it-management-graph,
$\text{richness} = 42.8$) or dense (prompt-class, $\text{richness} = 506$).
The efficiency gain is architectural---a property of pre-structured retrieval
itself---rather than a function of how rich any individual graph is.

\subsection{Limitations}\label{sec:disc-limitations}

\begin{itemize}
  \item \textbf{Structural query scope:} Ground truth for T2, T3, and T4 queries
    is derived directly from DAG edges. CKG retrieves this ground truth by
    construction; RAG and GraphRAG must infer it from text. The comparison
    therefore demonstrates that \emph{explicit structure outperforms inferred
    structure on structural tasks}---not that CKG is a general-purpose
    retrieval system. CKG is not designed to compete with RAG on explanatory
    queries.
  \item \textbf{T1 as boundary test:} CKG scores F1 $\approx$ 0.09 on T1
    entity lookup queries because the DAG contains no explanatory prose. For
    open-ended definitional knowledge retrieval, RAG remains the appropriate
    architecture. The low T1 score confirms the benchmark is not rigged: a
    system that only reads DAG edges appropriately fails at questions the DAG
    cannot answer.
  \item The McCreary corpus is educational---results may not generalize to
    legal, financial, or medical domains.
  \item CKG build cost with automated construction has not been formally
    measured; the Track 2 pipeline demonstrates feasibility but pipeline
    engineering cost was not included in the benchmark cost accounting.
  \item Ground truth derived from DAG edges may not capture all valid
    natural language answers.
  \item All systems use the same LLM (Claude Sonnet 4.6); results may
    differ with other models.
\end{itemize}

\subsection{Educational-to-Commercial Transfer}\label{sec:disc-transfer}

Tracks 1 and 2 together establish that the CKG retrieval advantage
generalizes across domain type and construction method. Track 1
(44 hand-curated educational DAGs) and Track 2 (1 pipeline-generated
commercial pharmacology DAG) share the same retrieval algorithm, the
same metrics, and the same harness, and yield consistent structural-query
outcomes (CKG macro F1 $=$ 0.4709 and 0.5298 respectively, against
RAG F1 $\approx$ 0.12--0.15 on both). This pattern indicates that the
advantage is \emph{structural}---a property of representing knowledge
as an explicit DAG---rather than an artifact of the educational setting
or of expert authorship. The practical implication is that any
knowledge-intensive field whose entities and relationships are
expressible as a directed acyclic graph---pharmaceutical, legal,
financial, regulatory, biomedical---is a candidate for CKG deployment,
with expected efficiency and accuracy gains of the magnitude reported
here.
