\section{The Economics of Learning Graph Generation}\label{sec:lg-economics}

A central limitation of any structure-first retrieval architecture is the
up-front cost of constructing the underlying knowledge structure. Section
\ref{sec:disc-limitations} notes that the benchmark's zero-build-cost
assumption for CKG is not strictly accurate: the learning graph itself must
be authored. Historically, this has been the decisive objection to structured
approaches---expert curation is slow, expensive, and does not scale.

This section addresses that objection directly. We first provide a formal
definition of a learning graph, then present a cost model for generating one
via an agentic workflow, and finally extrapolate the trajectory of generation
cost over the next 18--24 months.

\subsection{Formal Definition of a Learning Graph}\label{sec:lg-def}

\begin{definition}[Learning Graph]\label{def:learning-graph}
A \emph{learning graph} is a 4-tuple $G = (C, E, T, \tau)$ where:
\begin{itemize}
  \item $C = \{c_1, c_2, \ldots, c_n\}$ is a finite, non-empty set of
    \emph{concepts}, each a named unit of domain knowledge with a unique
    identifier and a human-readable label.
  \item $E \subseteq C \times C$ is a set of directed \emph{prerequisite
    edges}. An edge $(c_i, c_j) \in E$ asserts that concept $c_i$ must be
    understood before concept $c_j$ can be meaningfully taught or applied.
  \item $T = \{t_1, t_2, \ldots, t_k\}$ is a finite set of \emph{taxonomy
    categories} that partition the concept space into coarse groupings
    (e.g., \texttt{FOUND}, \texttt{CORE}, \texttt{ADV}).
  \item $\tau : C \rightarrow T$ is a total function assigning each concept
    to exactly one taxonomy category.
\end{itemize}
The pair $(C, E)$ must form a \emph{directed acyclic graph} (DAG): there
exists no sequence $c_{i_1}, c_{i_2}, \ldots, c_{i_m}, c_{i_1}$ such that
every consecutive pair is an edge in $E$. Acyclicity ensures a valid
teaching order exists (topological sort over $E$).
\end{definition}

The concrete serialization used throughout this paper and in the McCreary
Intelligent Textbook Corpus encodes $G$ as a four-column CSV:
\texttt{ConceptID}, \texttt{ConceptLabel}, \texttt{Dependencies},
\texttt{TaxonomyID}. The \texttt{Dependencies} column is a pipe-delimited
list of prerequisite \texttt{ConceptID} values, directly representing the
in-edges of each concept in $E$. This serialization is deliberately minimal:
a 200-concept learning graph typically fits in 6--12 KB of plain text.

Three properties follow directly from Definition~\ref{def:learning-graph}
and are load-bearing for the CKG architecture:
\begin{enumerate}
  \item \textbf{Finite, enumerable context.} Because $C$ is finite and the
    CSV serialization is compact, the entire graph fits in an LLM prompt
    for any realistic domain ($|C| \leq 1{,}000$).
  \item \textbf{Deterministic traversal.} Prerequisite chains are computed
    by BFS or DFS over $E$ with no inference required. Query answers for
    structural questions (T2, T3, T4) are functions, not predictions.
  \item \textbf{Closed vocabulary.} The set of valid concept labels is
    exactly $\{\mathrm{label}(c) : c \in C\}$. A retrieval system that
    returns only concepts in $C$ cannot hallucinate entities by construction.
\end{enumerate}

\subsection{Agentic Learning Graph Generation}\label{sec:lg-gen}

The McCreary corpus graphs used in this benchmark were produced with an
agentic workflow (the \texttt{/learning-graph-generator} skill running
under Claude Code) that decomposes generation into three stages:
\begin{enumerate}
  \item \textbf{Concept elicitation.} Given a course description, the
    agent proposes a concept set $C$ of a target size (default $n=200$),
    drawing on the model's domain knowledge and any supplied source
    materials.
  \item \textbf{Dependency assignment.} For each concept $c_j \in C$, the
    agent selects the subset of previously-enumerated concepts on which
    $c_j$ depends, populating $E$ incrementally.
  \item \textbf{Taxonomy assignment and validation.} The agent assigns
    $\tau(c)$ for each $c \in C$, then runs automated validation: cycle
    detection, orphan detection, dependency count distribution, and a
    quality score. Graphs scoring below threshold trigger a correction
    cycle.
\end{enumerate}

A subject-matter expert (SME) reviews the final graph for domain fidelity
and edits edges or labels as needed. In practice, SME review for a
200-concept graph takes 2--4 hours; the agentic generation itself
completes in minutes.

\subsection{Cost Model}\label{sec:lg-cost}

\paragraph{Measurement methodology.}
Rather than estimating generation cost from first principles, we measured
it directly. Each invocation of the \texttt{/learning-graph-generator}
skill is recorded by a Claude Code \texttt{PostToolUse} hook
(\texttt{track-skill-end.sh}) that writes a \texttt{skill-usage.jsonl}
event with the session identifier, project path, and wall-clock
duration. The full session transcript---every assistant message with a
complete \texttt{usage} object reporting \texttt{input\_tokens},
\texttt{output\_tokens}, \texttt{cache\_read\_input\_tokens}, and
\texttt{cache\_creation\_input\_tokens}---is retained by Claude Code at
\texttt{\textasciitilde/.claude/projects/\{slug\}/\{session\}.jsonl}. We
walk each transcript, sum token usage across all assistant messages from
the \texttt{/learning-graph-generator} skill-start timestamp to the next
skill-start event in the same session, and apply the per-model public
pricing for Claude Opus 4.6
(\$15/M input, \$75/M output, \$1.50/M cache read, \$18.75/M cache
write). The numbers in Table~\ref{tab:lg-cost-measured} are
\emph{measured}, not estimated. The measurement script and the raw CSV
are included in the repository at
\texttt{paper/figures/measure\_lg\_generation\_cost.py}.

Of 21 recorded \texttt{/learning-graph-generator} invocations, 9 had
surviving full session transcripts at measurement time; the remaining 12
sessions ran before the transcript-retention policy in use and cannot be
reconstructed precisely. All 9 measured sessions used Claude Opus 4.6 as
the generating model.

\begin{table}[htbp]
\centering
\small
\caption{Measured token consumption and cost for nine complete
\texttt{/learning-graph-generator} sessions, sorted by concept count.
Costs reflect Claude Opus 4.6 public API pricing at measurement
time.\protect\footnotemark{} Cached column sums cache-read and
cache-write tokens, which dominate total volume because the growing
graph is re-attached to each assistant turn.}
\label{tab:lg-cost-measured}
\input{figures/lg_generation_cost_table}
\end{table}
\footnotetext{The sessions reported here were generated under a Claude
Max subscription, a flat monthly fee with no per-token metering, so no
incremental dollars were paid for these specific runs. The costs in the
table are computed from the public pay-as-you-go API rates so that a
reader evaluating whether to adopt this workflow sees the figure they
would themselves be charged.}

\paragraph{Fitted cost model.}
We model generation cost as an affine function of concept count:
\begin{equation}\label{eq:lg-cost}
  \mathrm{Cost}(n) \approx \alpha + \beta \cdot n
\end{equation}
where $\alpha$ is the fixed-overhead cost and $\beta$ is the marginal
per-concept cost. A least-squares fit to the nine measured sessions
yields, for Claude Opus 4.6:
\begin{equation}\label{eq:lg-fit-opus}
  \mathrm{Cost}_{\mathrm{Opus\;4.6}}(n) \approx \$8.16 + \$0.019 \cdot n
\end{equation}
with $R^2 \approx 0.24$---the low coefficient of determination reflects
substantial session-to-session variance driven by the number of
validation and correction cycles a given run requires, which dominates
any smooth dependence on concept count. The mean measured cost was
\$13.94 across sessions averaging 311 concepts, with observed values
ranging from \$9.21 to \$21.38. We therefore prefer to report
Table~\ref{tab:lg-cost-measured} directly rather than relying on the
fit; Equation~\ref{eq:lg-fit-opus} is provided for
ballpark estimation only.

For cheaper model tiers, direct measurement is not yet available because
the McCreary corpus was generated on Opus. Applying the per-token price
ratio between Opus and Sonnet 4.6 (roughly 5$\times$) to the measured
Opus numbers yields a projected Sonnet cost of approximately \$3 for a
200-concept graph; re-running \texttt{/learning-graph-generator} on
Sonnet and measuring directly would tighten this estimate.

\paragraph{Interpretation.}
Even at the highest current model tier, a domain learning graph costs
between ten and twenty-five dollars of compute to generate---excluding
the subsequent SME review loop, which remains the dominant human-time
cost but is no longer the dominant economic bottleneck. These figures
are four to five orders of magnitude below the cost of traditional
expert curation (a team of domain experts working for weeks or months)
that has historically blocked structure-first retrieval from commercial
deployment.

\subsection{Projected Trajectory}\label{sec:lg-projection}

Two compounding trends point toward sharply lower generation cost over
the next 18--24 months:

\begin{enumerate}
  \item \textbf{Declining model pricing at constant capability.} Across
    the preceding two years, intelligence-per-dollar at the Anthropic
    frontier has approximately halved every 9--12 months, as smaller
    models match the capability of their predecessors. Applied to
    Equation~\ref{eq:lg-cost}, both $\alpha$ and $\beta$ scale
    proportionally with per-token pricing.
  \item \textbf{Skill-level efficiency gains.} The current generation
    workflow holds the full emerging graph in context, producing a
    super-linear accumulation term in $\beta$. Decomposing generation
    into cache-efficient passes (concepts first, dependencies batched,
    taxonomy last) collapses the accumulation and has been observed to
    reduce token consumption 2--3$\times$ at any given model tier.
\end{enumerate}

Combining these trends with the measured Opus 4.6 baseline of roughly
\$12 per 200-concept graph (Table~\ref{tab:lg-cost-measured}), we project
a plausible 2027 cost of \$1--\$2 per 200-concept graph when a
Haiku-class successor model is used with a cache-efficient rewrite of
the generation workflow. At that price point, domain graph generation
is effectively free relative to any downstream inference workload.

\subsection{Implications for the Build-Cost Objection}\label{sec:lg-implications}

The traditional objection to structure-first retrieval---\emph{``curating a
domain graph is prohibitively expensive''}---was empirically true for
two decades. It is no longer true. When a 200-concept domain graph can be
generated for dollars of compute and a handful of SME-review hours, the
decision calculus inverts: the question is no longer whether a domain can
afford a learning graph, but whether any high-query-volume domain can
afford \emph{not} to have one, given that per-query retrieval costs fall
by roughly an order of magnitude once the graph exists
(Section~\ref{sec:disc-ckg-wins}).

This shifts where the economic moat lies. Model compute is commoditizing;
the durable value in applying CKG to new domains is the
\emph{SME review loop}---ensuring the generated graph faithfully reflects
the expert consensus of the field---and the \emph{schema design} work
required to extend Definition~\ref{def:learning-graph} beyond
prerequisite-structured domains to domains with richer relation types
(Section~\ref{sec:disc-limitations}).
