\section{Conclusion}\label{sec:conclusion}

We presented the CKG Benchmark, a two-track comparison of Compact
Knowledge Graphs against RAG and GraphRAG across 44 hand-curated
educational domains (Track 1, 7,758 queries) and one pipeline-generated
commercial pharmacology domain (Track 2, 170 queries). Our key findings
are:

\begin{enumerate}
  \item \textbf{CKG achieves 4$\times$ higher macro F1 than RAG} on
    structural knowledge queries (\textbf{0.4709} vs.\ \textbf{0.1231}
    vs.\ \textbf{0.1200} for GraphRAG), using \textbf{11$\times$ fewer
    tokens per query} (269 vs.\ 2,982 vs.\ 3,450). The compound RDS
    advantage is \textbf{42$\times$} on Track 1.
  \item \textbf{Structural query performance is near-ceiling for
    aggregation}: T4 category aggregates reach F1 $=$ \textbf{0.964}
    for CKG vs.\ 0.286 for RAG and 0.054 for GraphRAG, because the
    answer is fully determined by the taxonomy column of the CSV.
  \item \textbf{CKG F1 increases with hop depth} (peaking at hop=3:
    \textbf{0.64}) while RAG degrades, confirming that deterministic
    edge traversal does not suffer the multi-hop retrieval recall
    failures inherent to passage-based systems.
  \item \textbf{T1 (entity lookup) is a documented negative control}:
    CKG scores \textbf{0.207} on explanatory queries, confirming the
    benchmark is not constructed to favour CKG universally. Both
    systems struggle on T1, which requires prose definitions absent
    from DAG structure.
  \item \textbf{RDS and Hop-Depth F1} are practical additions to IR
    evaluation that jointly measure quality and efficiency---metrics
    missing from existing benchmarks such as BEIR and RAGAS.
  \item \textbf{The CKG advantage transfers to a pipeline-generated
    commercial domain.} Track 2 builds a GLP-1/Obesity CKG
    programmatically from the ClinicalTrials.gov API, with no expert
    curation and no educational source material. The resulting domain
    achieves CKG macro F1 $=$ \textbf{0.5298}, exceeding the Track 1
    hand-curated average by 12.5\%, and preserves a \textbf{28$\times$}
    RDS advantage over RAG. Automated construction does not degrade
    retrieval quality.
\end{enumerate}

\paragraph{Limitations.}
Ground truth for T2--T4 is derived from the same DAG used by CKG, which
is a methodological constraint acknowledged in the design
(Section~\ref{sec:benchmark}). T1 performance on explanatory queries is
weak for all three systems and represents an open problem. Pipeline
engineering cost for automated CKG construction (Track 2) was not
included in the benchmark cost accounting; build-cost accounting across
construction strategies remains future work.

\paragraph{Future Work.}
\begin{itemize}
  \item Extend the automated construction pipeline to additional
    structured data sources (regulatory registries, legal codes,
    financial filings, product catalogues) to test the Track 2 transfer
    result at scale.
  \item Extend T5 cross-concept retrieval with deeper BFS path-finding;
    the current T5 F1 of 0.323 (Track 1) and 0.425 (Track 2) has a clear
    improvement trajectory.
  \item Hybrid architectures combining CKG structural precision with
    RAG's prose retrieval for T1 and T5 query types.
  \item Formal build-cost accounting that includes pipeline engineering,
    taxonomy-schema design, and source-API cost, enabling end-to-end
    CPCA comparison across all three architectures.
  \item Model-robustness experiments replicating the benchmark across
    additional LLM families.
\end{itemize}

\paragraph{Unified contribution.}
Track 1 validates the intelligent textbook format---hand-curated
learning-graph CSVs---as a benchmark-grade substrate for knowledge
retrieval at scale across 44 domains. Track 2 validates automated
construction of CKG domains from structured external data, showing the
retrieval advantage does not depend on manual authorship. Taken
together, the two tracks establish Compact Knowledge Graphs as a
distinct architecture category---one that is empirically superior to
RAG and GraphRAG on structural queries, an order of magnitude cheaper
in tokens, and ready for deployment in any domain whose knowledge can
be expressed as a directed acyclic graph.

\paragraph{Open Benchmark.}
The complete benchmark---corpus, queries, evaluation harness, and all
results---is released at
\url{https://huggingface.co/datasets/graphify-md/ckg-benchmark}
under CC BY 4.0 (data) and MIT (code). We invite the community to add
domains, systems, and metrics.
